### 2017-05-22 2017mergingmerges/section/04-Fusion.tex
Address reviewer comments

% -----------------------------------------------------------------------------
\subsection{Fusibility}
\label{s:FusionOrder}
When we fuse a pair of processes we commit to a particular interleaving of instructions from each process. When we have at least three processes to fuse, the choice of which two to handle first can determine whether this fused result can then be fused with the third process. Consider the following example, where @alt2@ pulls two elements from its first input stream, then two from its second, before pushing all four to its output.
\begin{code}
  alternates : S Nat -> S Nat -> S Nat -> S (Nat, Nat)
  alternates sInA sInB sInC
   = let  s1   = alt2 sInA sInB
          s2   = alt2 sInB sInC
          sOut = zip s1 s2
     in   sOut
\end{code}

In our current system, if we fuse the two @alt2@ processes together first, then try to fuse this result process with the downstream @zip@ process, the final fusion transform fails. This happens because the first fusion transform commits to a sequential instruction interleaving where two output values \emph{must} be pushed to stream @s1@ first, before pushing values to @s2@. On the other hand, @zip@ needs to pull a \emph{single} value from each of its inputs alternately.

Dynamically, if we were to execute the first fused result process, and the downstream @zip@ process concurrently, then the execution would deadlock. Statically, when we try to fuse the result process with the downstream @zip@ process the deadlock is discovered and fusion fails. Deadlock happens when neither process can advance to the next instruction, and in the fusion algorithm this manifests as the failure of the $tryStepPair$ function from Figure~\ref{fig:Fusion:Def:StepPair}. The $tryStepPair$ function determines which instruction from either process to execute next, and when execution is deadlocked there are none.

@@

In practice, the likelihood of fusion suceeding depends on the particular dataflow network being used, as well as the form of the processes in that network. For fusion of pipelines of standard combinators such as @map@, @fold@, @filter@, @scan@ and so on, fusion always succeeds. The process implementations of each of these combinators only pull one element at a time from their source streams, before pushing the result to the output stream, so there is no possiblity of deadlock. Deadlock can only happen when multiple streams fan-in to a process with multiple inputs, such as with @merge@.

>>>>

% -----------------------------------------------------------------------------
\subsection{Fusibility}
\label{s:FusionOrder}
When we fuse a pair of processes we commit to a particular interleaving of instructions from each process. When we have at least three processes to fuse, the choice of which two to handle first can determine whether this fused result can then be fused with the third process. Consider the following example, where @alt2@ pulls two elements from its first input stream, then two from its second, before pushing all four to its output.
\begin{code}
 alternates : S Nat -> S Nat -> S Nat -> S (Nat, Nat)
 alternates sInA sInB sInC
  = let  s1   = alt2 sInA sInB
         s2   = alt2 sInB sInC
         sOut = zip s1 s2
    in   sOut
\end{code}
If we fuse the two @alt2@ processes together first, then try to fuse this result process with the downstream @zip@ process, the final fusion transform fails. This happens because the first fusion transform commits to a sequential instruction interleaving where two output values \emph{must} be pushed to stream @s1@ first, before pushing values to @s2@. On the other hand, @zip@ needs to pull a \emph{single} value from each of its inputs alternately.

Dynamically, if we were to execute the first fused result process, and the downstream @zip@ process concurrently, then the execution would deadlock. Statically, when we try to fuse the result process with the downstream @zip@ process the deadlock is discovered and fusion fails. Deadlock happens when neither process can advance to the next instruction, and in the fusion algorithm this manifests as the failure of the $tryStepPair$ function from Figure~\ref{fig:Fusion:Def:StepPair}. The $tryStepPair$ function determines which instruction from either process to execute next, and when execution is deadlocked there are none. Fusability is an under-approximation for \emph{deadlock freedom} of the network.

@@

In practice, the likelihood of fusion suceeding depends on the particular dataflow network. For fusion of pipelines of standard combinators such as @map@, @fold@, @filter@, @scan@ and so on, fusion always succeeds. The process implementations of each of these combinators only pull one element at a time from their source streams, before pushing the result to the output stream, so there is no possiblity of deadlock. Deadlock can only happen when multiple streams fan-in to a process with multiple inputs, such as with @merge@.
